PANCHANGE_data_preprocessing

### 2020-08-11 PANCHANGE_data_preprocessing2020-07-15.Rmd
added first edgi read in chunk

@@

glad.employment.raw.id <- glad.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r glad gad cols}
glad.gad.raw <- readRDS(file = paste0(data.raw_path, "/glad/gad_glad.rds"))
dim(glad.gad.raw)
colnames(glad.gad.raw)
```

@@

@@

# Calulate total score
gad.scored_items_prepan <- scoreItems(keys = gad.items_key_prepan,
                                        items = glad.gad.raw.id[gad.items_prepan],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', 
                                        min = 0, 
                                        max = 3)
# Add in column to tibble
glad.gad.raw.id$gad.sum_score_prepan_raw <- gad.scored_items_prepan$scores

# Look at values
glad.gad.raw.id %>%
  descr(gad.sum_score_prepan_raw,
        stats = "common"
        )
```

@@

prepandemic
```{r gad prepan missingness}
glad.gad.raw.id <- glad.gad.raw.id %>%
  mutate(
    na_per_person_gad_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[gad.items_prepan])
      )
  )

glad.gad.raw.id %>%
  freq(na_per_person_gad_prepan,
       cumul = F)
```

```{r gad prepan sum score if missigng items set to NA}
glad.gad.raw.id <- glad.gad.raw.id %>%
  mutate(
    gad.sum_score_prepan =
      if_else(condition = na_per_person_gad_prepan > 0,
              true = NA_real_,
              false = gad.sum_score_prepan_raw,
              missing = NA_real_
              )
  )

glad.gad.raw.id %>%
  descr(gad.sum_score_prepan_raw, stats = "common")

glad.gad.raw.id %>%
  descr(gad.sum_score_prepan, stats = "common")
```

```{r glad gad head}
head(glad.gad.raw.id)
```

```{r glad gad summary}
summarytools::dfSummary(
  glad.gad.raw.id,
  graph.col = F,
  valid.col = F,
  labels.col = F)
```

>>>>

```{r nbr copy dem}
if(NBR == TRUE){
dem.raw.id <- nbr.dem.raw.id
}
```

@@

glad.employment.raw.id <- glad.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r glad gad cols}
if(GLAD == TRUE){
glad.gad.raw <- readRDS(file = paste0(data.raw_path, "/glad/gad_glad.rds"))
dim(glad.gad.raw)
colnames(glad.gad.raw)
}
```

@@

```{r nbr copy gad}
if(NBR == TRUE){
gad.raw.id <- glad.gad.raw.id
}
```

```{r edgi gad cols}
if(EDGI == TRUE){
  edgi.gad.raw <- readRDS(file = paste0(data.raw_path, "/edgi/gad_edgi.rds"))
  dim(edgi.gad.raw)
  colnames(edgi.gad.raw)
}
```

```{r edgi gad cleaning}
exclude_cols <- c("Sample",
                  "ID")

edgi.gad.raw.id <- edgi.gad.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.edgi = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    gad.feeling_nervous_anxious_or_on_edge_prepan = gad.feeling_nervous_anxious_or_on_edge,
    gad.control_worrying_stop_prepan = gad.control_worrying_stop,
    gad.worrying_too_much_about_different_things_prepan = gad.worrying_too_much_about_different_things,
    gad.trouble_relaxing_prepan = gad.trouble_relaxing,
    gad.sit_restless_hard_prepan = gad.sit_restless_hard,
    gad.becoming_easily_annoyed_or_irritable_prepan = gad.becoming_easily_annoyed_or_irritable,
    gad.awful_feeling_afraid_happen_prepan = gad.awful_feeling_afraid_happen
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(edgi.gad.raw.id)
# Inspect colnames
colnames(edgi.gad.raw.id)
#Differences
dim(edgi.gad.raw)[1]-dim(edgi.gad.raw.id)[1]
```

```{r nbr copy gad}
if(NBR == TRUE){
  gad.raw.id <- edgi.gad.raw.id
}
```

@@

# Calulate total score
gad.scored_items_prepan <- scoreItems(keys = gad.items_key_prepan,
                                        items = glad.gad.raw.id[gad.items_prepan],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', 
                                        min = 0, 
                                        max = 3)
# Add in column to tibble
GADraw.id$gad.sum_score_prepan_raw <- gad.scored_items_prepan$scores

# Look at values
GADraw.id %>%
  descr(gad.sum_score_prepan_raw,
        stats = "common"
        )
```

@@

prepandemic
```{r gad prepan missingness}
glad.gad.raw.id <- gad.raw.id %>%
  mutate(
    na_per_person_gad_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[gad.items_prepan])
      )
  )

gad.raw.id %>%
  freq(na_per_person_gad_prepan,
       cumul = F)
```

```{r gad prepan sum score if missigng items set to NA}
gad.raw.id <- gad.raw.id %>%
  mutate(
    gad.sum_score_prepan =
      if_else(condition = na_per_person_gad_prepan > 0,
              true = NA_real_,
              false = gad.sum_score_prepan_raw,
              missing = NA_real_
              )
  )

gad.raw.id %>%
  descr(gad.sum_score_prepan_raw, stats = "common")

gad.raw.id %>%
  descr(gad.sum_score_prepan, stats = "common")
```

```{r glad gad head}
head(gad.raw.id)
```

```{r glad gad summary}
summarytools::dfSummary(
  gad.raw.id,
  graph.col = F,
  valid.col = F,
  labels.col = F)
```
